---
title: "Texas County COVID Positive Test Rate Trend"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    social: menu
    source_code: https://github.com/mychan24/covid_dallas_trend
runtime: shiny

---
<style>                     
.navbar {
  background-color:#009a9a;
  border-color:black;
}
.navbar-brand {
color:white!important;
}
</style>         

```{r global, include=FALSE}
library(shiny)
library(flexdashboard)
library(tidyverse)
library(patchwork)
library(readxl)
library(plotly)

### Set Theme for ggplot ###
theme_set(theme_bw() +
            theme(strip.text = element_text(size=14, face="plain"),
              axis.title = element_text(size=14, face=("plain")),
              plot.title = element_text(size=16, face=("plain")))
)

theme_mobile <- theme_bw() +
              theme(strip.text = element_text(size=8, face="plain"),
              axis.title = element_text(size=8, face=("plain")),
              plot.title = element_text(size=10, face=("plain")))

# --- Download data from dshs.texas.gov --- #

download.file(url = "https://dshs.texas.gov/coronavirus/TexasCOVID19DailyCountyCaseCountData.xlsx", destfile = "./TexasCOVID19DailyCountyCaseCountData.xlsx", mode="wb")

download.file(url = "https://dshs.texas.gov/coronavirus/TexasCOVID-19CumulativeTestsOverTimebyCounty.xlsx",
              destfile = "./COVID-19CumulativeTestTotalsbyCounty.xlsx", mode="wb")


# --- Load Case and Test Data --- #
case <- read_excel(path = "./TexasCOVID19DailyCountyCaseCountData.xlsx",
                   skip =2, n_max = 258)

## remove extra spaces
names(case)[3:ncol(case)] <- gsub("\r", "", names(case)[3:ncol(case)])
names(case)[3:ncol(case)] <- gsub("\n", "", names(case)[3:ncol(case)])
names(case)[3:ncol(case)] <- gsub("Cases", "", names(case)[3:ncol(case)])
names(case)[3:ncol(case)] <- gsub(" ", "", names(case)[3:ncol(case)])

t <- read_excel(path = "./COVID-19CumulativeTestTotalsbyCounty.xlsx",
                skip =1, n_max = 258)

file.remove("./TexasCOVID19DailyCountyCaseCountData.xlsx")
file.remove("./COVID-19CumulativeTestTotalsbyCounty.xlsx")

names(t)[2:ncol(t)] <- sub("Tests Through ", "", names(t)[2:ncol(t)])
# --- Texas county data cleaning and calculating positive test rate (ptr) --- #
counties <- t$County[t[,ncol(t)]>3000]
counties <- counties[!is.element(counties, c("Unknown", "Pending Assignments", "TOTAL"))]

df_list <- list()
for(cc in 1:length(counties)){
  ## Dallas Positive Rate data
  df <- data.frame(Date=names(case)[(which(names(case)=="04-21")+1):ncol(case)],
                   Total.Test.Performed = as.numeric(t[t$County==counties[cc],2:ncol(t)]),
                   Total.Cases = as.numeric(case[which(case$`County Name`==counties[cc]),(which(names(case)=="04-21")+1):ncol(case)-1]))
  
  df$Positive.Rate <- df$Total.Cases/df$Total.Test.Performed*100
  df$Date <- as.Date(df$Date, "%m-%d")
  
  # --- Calculate New-case per day &  Positive Rate --- #
  ## Calculate new-test each day
  df$New.Case.Reported <- NA
  df$New.Test.Reported <- NA
  for(i in 2:nrow(df)){
    df$New.Case.Reported[i] <- df$Total.Cases[i] - df$Total.Cases[i-1]
    df$New.Test.Reported[i] <- df$Total.Test.Performed[i] - df$Total.Test.Performed[i-1]
  }
  
  ## Smooth the per-day data over 7 days
  df$New.Case.Reported_7davg <- NA
  df$New.Test.Reported_7davg <- NA
  for(i in 7:nrow(df)){
    df$New.Case.Reported_7davg[i] <- mean(df$New.Case.Reported[(i-6):i], na.rm=T)
    df$New.Test.Reported_7davg[i] <- mean(df$New.Test.Reported[(i-6):i], na.rm=T)
  }
  
  df$New.Case.Reported_7davg[df$New.Case.Reported_7davg < 0] <- NA # remove < 0 Test per day (due mis-reporting)
  df$New.Test.Reported_7davg[df$New.Test.Reported_7davg < 0] <- NA # remove < 0 Test per day (due mis-reporting)
  
  df$Positive.Rate.Per.Day <- df$New.Case.Reported/df$New.Test.Reported *100
  df$Positive.Rate.Per.Day_7davg <- df$New.Case.Reported_7davg/df$New.Test.Reported_7davg *100
  
  df$Positive.Rate.Per.Day_7davg[df$Positive.Rate.Per.Day_7davg > 50] <- NA # remove error due to Tests reported

  
  ## Label a positive rate data point as levels
  df$Positive.Rate.Level <- NA
  df$Positive.Rate.Level[df$Positive.Rate.Per.Day_7davg<3] <- "Low"
  df$Positive.Rate.Level[df$Positive.Rate.Per.Day_7davg>=3 & df$Positive.Rate.Per.Day_7davg < 10] <- "Medium"
  df$Positive.Rate.Level[df$Positive.Rate.Per.Day_7davg>=10 & df$Positive.Rate.Per.Day_7davg < 20] <- "High"
  df$Positive.Rate.Level[df$Positive.Rate.Per.Day_7davg>=20] <- "Critical"
  df$Positive.Rate.Level <- factor(df$Positive.Rate.Level, levels=c("Low","Medium","High","Critical"))
  
  df_list[[cc]] <- df
  names(df_list)[[cc]] <- counties[[cc]]
}


# --- Texas data --- #
df_texas <- data.frame(Date = names(case)[(which(names(case)=="04-21")+1):ncol(case)],
                       Total.Test.Performed = as.numeric(t[t$County=="TOTAL",2:ncol(t)]),
                       Total.Cases = as.numeric(case[which(case$`County Name`=="Total"), (which(names(case)=="04-21")+1):ncol(case)-1]))

df_texas$Positive.Rate <- df_texas$Total.Cases/df_texas$Total.Test.Performed*100
df_texas$Date <- as.Date(df_texas$Date, "%m-%d")

df_texas$New.Case.Reported <- NA
df_texas$New.Test.Reported <- NA
for(i in 2:nrow(df_texas)){
  df_texas$New.Case.Reported[i] <- df_texas$Total.Cases[i] - df_texas$Total.Cases[i-1]
  df_texas$New.Test.Reported[i] <- df_texas$Total.Test.Performed[i] - df_texas$Total.Test.Performed[i-1]
}

## Smooth the per-day data over 5 days
df_texas$New.Case.Reported_7davg <- NA
df_texas$New.Test.Reported_7davg <- NA
for(i in 7:nrow(df_texas)){
  df_texas$New.Case.Reported_7davg[i] <- mean(df_texas$New.Case.Reported[(i-6):i], na.rm=T)
  df_texas$New.Test.Reported_7davg[i] <- mean(df_texas$New.Test.Reported[(i-6):i], na.rm=T)
}

df_texas$Positive.Rate.Per.Day <- df_texas$New.Case.Reported/df_texas$New.Test.Reported *100
df_texas$Positive.Rate.Per.Day_7davg <- df_texas$New.Case.Reported_7davg/df_texas$New.Test.Reported_7davg *100

## Label a positive rate data point as levels
df_texas$Positive.Rate.Level <- NA
df_texas$Positive.Rate.Level[df_texas$Positive.Rate.Per.Day_7davg<3] <- "Low"
df_texas$Positive.Rate.Level[df_texas$Positive.Rate.Per.Day_7davg>=3 & df_texas$Positive.Rate.Per.Day_7davg < 10] <- "Medium"
df_texas$Positive.Rate.Level[df_texas$Positive.Rate.Per.Day_7davg>=10 & df_texas$Positive.Rate.Per.Day_7davg < 20] <- "High"
df_texas$Positive.Rate.Level[df_texas$Positive.Rate.Per.Day_7davg>=20] <- "Critical"
df_texas$Positive.Rate.Level <- factor(df_texas$Positive.Rate.Level, levels=c("Low","Medium","High","Critical"))

```

Sidebar {.sidebar}
======================================================================

<font size="-2"><p> __Disclaimer:__ These data were collected different (i) geographical locations, (ii) reported by different entities, and (iii) based only on 'individuals that got tested'. The trends should be interpreted or compared with great caution. These trends are only as accurate as the underlying "reported" cases/tests.</p></font>

```{r sidebar}
# Define inputs
selectInput('county_name', label = 'Select a county:', choices = counties, selected = "Dallas")

```

Select __State Trend tab__ to see state-level trend data.

__Positive test rate__ is one of the frequently quoted metric regarding COVID trend. Most websites/reports have available state-level  data. This dashboard makes available county-level data for counties with > 3000 COVID tests performed to date. 

__Missing data__ is present in the raw data at times, where no 'additional' cases/tests were reported for a day. Those data-points were removed from the dashboard.

__Download data:__ [Texas DSHS data](https://dshs.texas.gov/coronavirus/additionaldata/)

<br><br>
Application author: [Micaela Chan](https://micaelachan.com)

<font size="-3">
<p>This program is free software: you can redistribute it and/or modify it under the terms of the GNU General Public License as published by the Free Software Foundation, either version 3 of the License, or (at your option) any later version. This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License for more details. You should have received a copy of the GNU General Public License along with this program. If not, see here.</p>
<p>This website is licensed under a Creative Commons Attribution-ShareAlike 4.0 International License.</p>
</font>


County Trend
=======================================================================

Row 
-----------------------------------------------------------------------
### County-level COVID Positive Test Rate Over Time

```{r fig_rate, fig.width=14, fig.height=4}
renderPlot(expr = {
df <- df_list[[input$county_name]]

most_recent_ptr <- sprintf("%s Positive Rate: %.1f%%", format(as.Date(df$Date[nrow(df)]), "%B %d, %Y"), df$Positive.Rate.Per.Day_7davg[nrow(df)])

  g1 <- df %>%
    ggplot(aes(x=Date, y=Positive.Rate.Per.Day_7davg)) +
    geom_point(aes(color=Positive.Rate.Level)) +
    # Re-open vertical lines
    geom_vline(xintercept = as.Date("2020-04-30"), color="firebrick", linetype="dashed", alpha=0.8) +
    annotate(geom = "text", x = as.Date("2020-04-28"), color="firebrick",  
             y = 28, label = "Re-open") +
    # Zone levels horizontal lines
    geom_hline(yintercept = 3, color="grey", linetype="dashed") +
    geom_hline(yintercept = 10, color="grey", linetype="dashed") +
    geom_hline(yintercept = 20, color="grey", linetype="dashed") +
    annotate(geom = "text", x = as.Date("2020-04-22"), y = 1, label = "Low", color="darkgreen") +
    annotate(geom = "text", x = as.Date("2020-04-23"), y = 6, label = "Medium", color="orange") +
    annotate(geom = "text", x = as.Date("2020-04-22"), y = 15, label = "High", color="#bb5516") +
    annotate(geom = "text", x = as.Date("2020-04-22"), y = 25, label = "Critical", color="red") +
    # text for ptr
    annotate(geom = "text", x= as.Date(df$Date[nrow(df)-8]), y=max(df$Positive.Rate.Per.Day_7davg,na.rm=T)+2, label=most_recent_ptr, color="grey20", size=4.5) + 
    # plot settings
    scale_color_manual(values = c("Low" = "darkgreen", "Medium" = "orange", "High" = "#bb5516", "Critical" = "red")) +
    scale_x_date(name = 'Date', date_breaks = '2 days', date_labels = '%m-%d', 
                 limits = c(as.Date("2020-04-22"), as.Date(df$Date[nrow(df)+1]))) +
    scale_y_continuous(labels = function(x) paste0(x, "%")) +
    ylab("Positive Rate Per Day(%)\n(New Case/New Test Per Day)") +
    ggtitle(label = "Est. Positive Rate Per Day Over Time (7-day average)",
            subtitle = sprintf("%s County", input$county_name)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.subtitle = element_text(size=14, color = "grey20", face = "bold"), 
          legend.position = "none") 
  
  g1 
})
```


Row
-----------------------------------------------------------------------
### County-level New COVID Confirmed Cases & Tests Over Time

```{r fig_casetest, fig.width=14, fig.height=6}

renderPlot({
df <- df_list[[input$county_name]]

g2 <- df %>%
    ggplot(aes(x=Date, y=New.Case.Reported_7davg)) +
    geom_bar(stat = "identity", fill="#cc5500") +
    # Re-open vertical lines
    geom_vline(xintercept = as.Date("2020-04-30", "%Y-%m-%d"), color="firebrick", linetype="dashed", alpha=0.8) +
    annotate(geom = "text", x = as.Date("2020-04-28", "%Y-%m-%d"),color="firebrick",  
             y = max(df$New.Case.Reported_7davg,na.rm=T)-max(df$New.Case.Reported_7davg,na.rm=T)/4, label = "Re-open") +
    # plot settings
    scale_x_date(name = 'Date', date_breaks = '2 days', date_labels = '%m-%d', 
                 limits = c(as.Date("2020-04-22", "%Y-%m-%d"), as.Date(df$Date[nrow(df)+1],"%Y-%m-%d"))) +
    ylab("New Cases Per Day\n(7-day average)") +
    ggtitle(label = "New Cases Per Day Over Time (7-day average)", 
            subtitle = sprintf("%s County", input$county_name)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.subtitle = element_text(size=14, color = "grey20", face = "bold"))

g3 <- df %>%
    ggplot(aes(x=Date, y=New.Test.Reported_7davg)) +
    geom_bar(stat = "identity", fill="#009a9a") +
    # Re-open vertical lines
    geom_vline(xintercept = as.Date("2020-04-30", "%Y-%m-%d"), color="firebrick", linetype="dashed", alpha=0.8) +
    annotate(geom = "text", x = as.Date("2020-04-28", "%Y-%m-%d"),color="firebrick",  
             y = max(df$New.Test.Reported_7davg,na.rm=T)-max(df$New.Test.Reported_7davg,na.rm=T)/4, label = "Re-open") +
    # plot settings
    scale_x_date(name = 'Date', date_breaks = '2 days', date_labels = '%m-%d', 
                 limits = c(as.Date("2020-04-22", "%Y-%m-%d"), as.Date(df$Date[nrow(df)+1],"%Y-%m-%d"))) +
    ylab("New Test Per Day\n(7-day average)") +
    ggtitle(label = "New Tests Per Day Over Time (7-day average)", 
            subtitle = sprintf("%s County", input$county_name)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.subtitle = element_text(size=14, color = "grey20", face = "bold"))
  
g2/g3
})

```

Texas State Trend
=======================================================================

Row 
-----------------------------------------------------------------------
### Texas State-level COVID Positive Test Rate Over Time


```{r fig_texas_rate, fig.width=14, fig.height=4}
renderPlot(expr = {

most_recent_ptr <- sprintf("%s Positive Rate: %.1f%%", format(as.Date(df_texas$Date[nrow(df_texas)]), "%B %d, %Y"), df_texas$Positive.Rate.Per.Day_7davg[nrow(df_texas)])

  g4 <- df_texas %>%
    ggplot(aes(x=Date, y=Positive.Rate.Per.Day_7davg)) +
    geom_point(aes(color=Positive.Rate.Level)) +
    # Re-open vertical lines
    geom_vline(xintercept = as.Date("2020-04-30"), color="firebrick", linetype="dashed", alpha=0.8) +
    annotate(geom = "text", x = as.Date("2020-04-28"), color="firebrick",  
             y = 28, label = "Re-open") +
    # Zone levels horizontal lines
    geom_hline(yintercept = 3, color="grey", linetype="dashed") +
    geom_hline(yintercept = 10, color="grey", linetype="dashed") +
    geom_hline(yintercept = 20, color="grey", linetype="dashed") +
    annotate(geom = "text", x = as.Date("2020-04-22"), y = 1, label = "Low", color="darkgreen") +
    annotate(geom = "text", x = as.Date("2020-04-23"), y = 6, label = "Medium", color="orange") +
    annotate(geom = "text", x = as.Date("2020-04-22"), y = 15, label = "High", color="#bb5516") +
    annotate(geom = "text", x = as.Date("2020-04-22"), y = 25, label = "Critical", color="red") +
    # text for ptr
    annotate(geom = "text", x= as.Date(df_texas$Date[nrow(df)-8]), y=30, label=most_recent_ptr, color="grey20", size=4.5) + 
    # plot settings
    scale_color_manual(values = c("Low" = "darkgreen", "Medium" = "orange", "High" = "#bb5516", "Critical" = "red")) +
    scale_x_date(name = 'Date', date_breaks = '2 days', date_labels = '%m-%d', 
                 limits = c(as.Date("2020-04-22"), as.Date(df_texas$Date[nrow(df)+1]))) +
    scale_y_continuous(labels = function(x) paste0(x, "%")) +
    ylab("Positive Rate Per Day(%)\n(New Case/New Test Per Day)") +
    ggtitle(label = "Est. Positive Rate Per Day Over Time (7-day average)") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.subtitle = element_text(size=14, color = "grey20", face = "bold"), 
          legend.position = "none") 
  
  g4
})
```


Row
-----------------------------------------------------------------------
### Texas State-level New COVID Confirmed Cases & Tests Over Time

```{r fig_texas_casetest, fig.width=14, fig.height=6}

renderPlot({
  
g5 <- df_texas %>%
    ggplot(aes(x=Date, y=New.Case.Reported_7davg)) +
    geom_bar(stat = "identity", fill="#cc5500") +
    # Re-open vertical lines
    geom_vline(xintercept = as.Date("2020-04-30", "%Y-%m-%d"), color="firebrick", linetype="dashed", alpha=0.8) +
    annotate(geom = "text", x = as.Date("2020-04-28", "%Y-%m-%d"),color="firebrick",  
             y = max(df_texas$New.Case.Reported_7davg,na.rm=T)-max(df_texas$New.Case.Reported_7davg,na.rm=T)/4, label = "Re-open") +
    # plot settings
    scale_x_date(name = 'Date', date_breaks = '2 days', date_labels = '%m-%d', 
                 limits = c(as.Date("2020-04-22", "%Y-%m-%d"), as.Date(df_texas$Date[nrow(df)+1],"%Y-%m-%d"))) +
    ylab("New Cases Per Day\n(7-day average)") +
    ggtitle(label = "New Cases Per Day Over Time (7-day average)") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.subtitle = element_text(size=14, color = "grey20", face = "bold"))

g6 <- df_texas %>%
    ggplot(aes(x=Date, y=New.Test.Reported_7davg)) +
    geom_bar(stat = "identity", fill="#009a9a") +
    # Re-open vertical lines
    geom_vline(xintercept = as.Date("2020-04-30", "%Y-%m-%d"), color="firebrick", linetype="dashed", alpha=0.8) +
    annotate(geom = "text", x = as.Date("2020-04-28", "%Y-%m-%d"),color="firebrick",  
             y = max(df_texas$New.Test.Reported_7davg,na.rm=T)-max(df_texas$New.Test.Reported_7davg,na.rm=T)/4, label = "Re-open") +
    # plot settings
    scale_x_date(name = 'Date', date_breaks = '2 days', date_labels = '%m-%d', 
                 limits = c(as.Date("2020-04-22", "%Y-%m-%d"), as.Date(df_texas$Date[nrow(df)+1],"%Y-%m-%d"))) +
    ylab("New Test Per Day\n(7-day average)") +
    ggtitle(label = "New Tests Per Day Over Time (7-day average)") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.subtitle = element_text(size=14, color = "grey20", face = "bold"))
  
g5/g6
})

```