---
title: "Dallas County and Texas's COVID-19: Trend in New Cases and Positive-Test-Rate"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    social: menu
    source_code: https://github.com/mychan24/covid_dallas_trend
    includes:
      after_body: footer.html
runtime: shiny

---
<style>                     
.navbar {
  background-color:#009a9a;
  border-color:black;
}
.navbar-brand {
color:white!important;
}
</style>         

```{r setup, include=FALSE}
library(shiny)
library(flexdashboard)
library(tidyverse)
library(patchwork)
library(readxl)

### Set Theme for ggplot ###
theme_set(theme_bw() +
            theme(strip.text = element_text(size=14, face="plain"),
              axis.title = element_text(size=14, face=("plain")),
              plot.title = element_text(size=16, face=("plain")))
)

theme_mobile <- theme_bw() +
              theme(strip.text = element_text(size=8, face="plain"),
              axis.title = element_text(size=8, face=("plain")),
              plot.title = element_text(size=10, face=("plain")))

# --- Download data from dshs.texas.gov --- #

download.file(url = "https://dshs.texas.gov/coronavirus/TexasCOVID19DailyCountyCaseCountData.xlsx", destfile = "../data/TexasCOVID19DailyCountyCaseCountData.xlsx", mode="wb")

download.file(url = "https://dshs.texas.gov/coronavirus/TexasCOVID-19CumulativeTestsOverTimebyCounty.xlsx",
              destfile = "../data/COVID-19CumulativeTestTotalsbyCounty.xlsx", mode="wb")


# --- Load Case and Test Data --- #
case <- read_excel(path = "../data/TexasCOVID19DailyCountyCaseCountData.xlsx",
                   skip =2, n_max = 258)

## remove extra spaces
names(case)[3:ncol(case)] <- gsub("\r", "", names(case)[3:ncol(case)])
names(case)[3:ncol(case)] <- gsub("\n", "", names(case)[3:ncol(case)])
names(case)[3:ncol(case)] <- gsub("Cases", "", names(case)[3:ncol(case)])
names(case)[3:ncol(case)] <- gsub(" ", "", names(case)[3:ncol(case)])

t <- read_excel(path = "../data/COVID-19CumulativeTestTotalsbyCounty.xlsx",
                skip =1, n_max = 258)

names(t)[2:ncol(t)] <- sub("Tests Through ", "", names(t)[2:ncol(t)])
# --- Texas county data cleaning and calculating positive test rate (ptr) --- #
counties <- t$County[t[,ncol(t)]>3000]
counties <- counties[!is.element(counties, c("Unknown", "Pending Assignments", "TOTAL"))]

df_list <- list()
for(cc in 1:length(counties)){
  ## Dallas Positive Rate data
  df <- data.frame(Date=names(case)[(which(names(case)=="04-21")+1):ncol(case)],
                   Total.Test.Performed = as.numeric(t[t$County==counties[cc],2:ncol(t)]),
                   Total.Cases = as.numeric(case[which(case$`County Name`==counties[cc]),(which(names(case)=="04-21")+1):ncol(case)-1]))
  
  df$Positive.Rate <- df$Total.Cases/df$Total.Test.Performed*100
  df$Date <- as.Date(df$Date, "%m-%d")
  
  # --- Calculate New-case per day &  Positive Rate --- #
  ## Calculate new-test each day
  df$New.Case.Reported <- NA
  df$New.Test.Reported <- NA
  for(i in 2:nrow(df)){
    df$New.Case.Reported[i] <- df$Total.Cases[i] - df$Total.Cases[i-1]
    df$New.Test.Reported[i] <- df$Total.Test.Performed[i] - df$Total.Test.Performed[i-1]
  }
  
  ## Smooth the per-day data over 7 days
  df$New.Case.Reported_7davg <- NA
  df$New.Test.Reported_7davg <- NA
  for(i in 3:nrow(df)){
    df$New.Case.Reported_7davg[i] <- mean(df$New.Case.Reported[(i-3):(i+3)], na.rm=T)
    df$New.Test.Reported_7davg[i] <- mean(df$New.Test.Reported[(i-3):(i+3)], na.rm=T)
  }
  
  df$Positive.Rate.Per.Day <- df$New.Case.Reported/df$New.Test.Reported *100
  df$Positive.Rate.Per.Day_7davg <- df$New.Case.Reported_7davg/df$New.Test.Reported_7davg *100
  
  ## Label a positive rate data point as levels
  df$Positive.Rate.Level <- NA
  df$Positive.Rate.Level[df$Positive.Rate.Per.Day_7davg<3] <- "Low"
  df$Positive.Rate.Level[df$Positive.Rate.Per.Day_7davg>=3 & df$Positive.Rate.Per.Day_7davg < 10] <- "Medium"
  df$Positive.Rate.Level[df$Positive.Rate.Per.Day_7davg>=10 & df$Positive.Rate.Per.Day_7davg < 20] <- "High"
  df$Positive.Rate.Level[df$Positive.Rate.Per.Day_7davg>=20] <- "Critical"
  df$Positive.Rate.Level <- factor(df$Positive.Rate.Level, levels=c("Low","Medium","High","Critical"))
  
  df_list[[cc]] <- df
  names(df_list)[[cc]] <- counties[[cc]]
}


# --- Texas state-level data cleaning --- #
df_texas <- data.frame(Date = names(case)[(which(names(case)=="04-21")+1):ncol(case)],
                       Total.Test.Performed = as.numeric(t[t$County=="TOTAL",2:ncol(t)]),
                       Total.Cases = as.numeric(case[which(case$`County Name`=="Total"), (which(names(case)=="04-21")+1):ncol(case)-1]))

df_texas$Positive.Rate <- df_texas$Total.Cases/df_texas$Total.Test.Performed*100
df_texas$Date <- as.Date(df_texas$Date, "%m-%d")

df_texas$New.Case.Reported <- NA
df_texas$New.Test.Reported <- NA
for(i in 2:nrow(df_texas)){
  df_texas$New.Case.Reported[i] <- df_texas$Total.Cases[i] - df_texas$Total.Cases[i-1]
  df_texas$New.Test.Reported[i] <- df_texas$Total.Test.Performed[i] - df_texas$Total.Test.Performed[i-1]
}

## Smooth the per-day data over 5 days
df_texas$New.Case.Reported_7davg <- NA
df_texas$New.Test.Reported_7davg <- NA
for(i in 3:nrow(df_texas)){
  df_texas$New.Case.Reported_7davg[i] <- mean(df_texas$New.Case.Reported[(i-3):(i+3)], na.rm=T)
  df_texas$New.Test.Reported_7davg[i] <- mean(df_texas$New.Test.Reported[(i-3):(i+3)], na.rm=T)
}

df_texas$Positive.Rate.Per.Day <- df_texas$New.Case.Reported/df_texas$New.Test.Reported *100
df_texas$Positive.Rate.Per.Day_7davg <- df_texas$New.Case.Reported_7davg/df_texas$New.Test.Reported_7davg *100

## Label a positive rate data point as levels
df_texas$Positive.Rate.Level <- NA
df_texas$Positive.Rate.Level[df_texas$Positive.Rate.Per.Day_7davg<3] <- "Low"
df_texas$Positive.Rate.Level[df_texas$Positive.Rate.Per.Day_7davg>=3 & df_texas$Positive.Rate.Per.Day_7davg < 10] <- "Medium"
df_texas$Positive.Rate.Level[df_texas$Positive.Rate.Per.Day_7davg>=10 & df_texas$Positive.Rate.Per.Day_7davg < 20] <- "High"
df_texas$Positive.Rate.Level[df_texas$Positive.Rate.Per.Day_7davg>=20] <- "Critical"
df_texas$Positive.Rate.Level <- factor(df_texas$Positive.Rate.Level, levels=c("Low","Medium","High","Critical"))

```

Sidebar {.sidebar}
======================================================================

```{r}
# Define inputs
selectInput('county_name', label = 'Select a county:', choices = counties, selected = "Dallas")

```

Select a __County__ to see the trend of new COVID-19 tests and postive test rate over time. Only counties with more than 3000 total COVID tests are available. 

__Positive test rate__ is one of the frequently quoted metric regarding COVID trend. Most websites/reports have available state-level  data. This dashboard makes available county-level data for counties with > 3000 COVID tests performed to date. 

Application author: [Micaela Chan](https://micaelachan.com)



COVID-19 Texas County Testing Trend
=======================================================================

Row 
-----------------------------------------------------------------------
### County-level New COVID Positive Test-Rate Over Time

```{r, fig.width=14, fig.height=4}

renderPlot({
df <- df_list[[input$county_name]]
    
  g1 <- df %>%
    ggplot(aes(x=Date, y=Positive.Rate.Per.Day_7davg)) +
    geom_point(aes(color=Positive.Rate.Level)) +
    # Re-open vertical lines
    geom_vline(xintercept = as.Date("2020-04-30", "%Y-%m-%d"), color="firebrick", linetype="dashed", alpha=0.8) +
    annotate(geom = "text", x = as.Date("2020-04-28", "%Y-%m-%d"), color="firebrick",  
             y = 28, label = "Re-open") +
    # Zone levels horizontal lines
    geom_hline(yintercept = 3, color="grey", linetype="dashed") +
    geom_hline(yintercept = 10, color="grey", linetype="dashed") +
    geom_hline(yintercept = 20, color="grey", linetype="dashed") +
    annotate(geom = "text", x = as.Date("2020-04-22", "%Y-%m-%d"), y = 1, label = "Low", color="darkgreen") +
    annotate(geom = "text", x = as.Date("2020-04-23", "%Y-%m-%d"), y = 6, label = "Medium", color="orange") +
    annotate(geom = "text", x = as.Date("2020-04-22", "%Y-%m-%d"), y = 15, label = "High", color="#bb5516") +
    annotate(geom = "text", x = as.Date("2020-04-22", "%Y-%m-%d"), y = 25, label = "Critical", color="red") +
    # plot settings
    scale_color_manual(values = c("Low" = "darkgreen", "Medium" = "orange", "High" = "#bb5516", "Critical" = "red")) +
    scale_x_date(name = 'Date', date_breaks = '2 days', date_labels = '%m-%d', 
                 limits = c(as.Date("2020-04-22", "%Y-%m-%d"), as.Date(df$Date[nrow(df)+1],"%Y-%m-%d"))) +
    scale_y_continuous(labels = function(x) paste0(x, "%")) +
    ylab("Positive Rate Per Day(%)\n(New Case/New Test Per Day)") +
    ggtitle(label = "Est. Positive Rate Per Day Over Time (7-day average)",
            subtitle = sprintf("%s County", input$county_name)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.subtitle = element_text(size=14, color = "grey20", face = "bold"), 
          legend.position = "none") 
  g1
})

```


Row
-----------------------------------------------------------------------
### County-level New COVID Confirmed Cases & Tests Over Time

```{r, fig.width=14, fig.height=6}

renderPlot({
df <- df_list[[input$county_name]]

g2 <- df %>%
    ggplot(aes(x=Date, y=New.Case.Reported_7davg)) +
    geom_bar(stat = "identity", fill="#cc5500") +
    # Re-open vertical lines
    geom_vline(xintercept = as.Date("2020-04-30", "%Y-%m-%d"), color="firebrick", linetype="dashed", alpha=0.8) +
    annotate(geom = "text", x = as.Date("2020-04-28", "%Y-%m-%d"),color="firebrick",  
             y = max(df$New.Case.Reported_7davg,na.rm=T)-max(df$New.Case.Reported_7davg,na.rm=T)/4, label = "Re-open") +
    # plot settings
    scale_x_date(name = 'Date', date_breaks = '2 days', date_labels = '%m-%d', 
                 limits = c(as.Date("2020-04-22", "%Y-%m-%d"), as.Date(df$Date[nrow(df)+1],"%Y-%m-%d"))) +
    ylab("New Cases Per Day\n(7-day average)") +
    ggtitle(label = "New Cases Per Day Over Time (7-day average)", 
            subtitle = sprintf("%s County", input$county_name)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.subtitle = element_text(size=14, color = "grey20", face = "bold"))

g3 <- df %>%
    ggplot(aes(x=Date, y=New.Test.Reported_7davg)) +
    geom_bar(stat = "identity", fill="#009a9a") +
    # Re-open vertical lines
    geom_vline(xintercept = as.Date("2020-04-30", "%Y-%m-%d"), color="firebrick", linetype="dashed", alpha=0.8) +
    annotate(geom = "text", x = as.Date("2020-04-28", "%Y-%m-%d"),color="firebrick",  
             y = max(df$New.Test.Reported_7davg,na.rm=T)-max(df$New.Test.Reported_7davg,na.rm=T)/4, label = "Re-open") +
    # plot settings
    scale_x_date(name = 'Date', date_breaks = '2 days', date_labels = '%m-%d', 
                 limits = c(as.Date("2020-04-22", "%Y-%m-%d"), as.Date(df$Date[nrow(df)+1],"%Y-%m-%d"))) +
    ylab("New Test Per Day\n(7-day average)") +
    ggtitle(label = "New Case/Tests Per Day Over Time (7-day average)", 
            subtitle = sprintf("%s County", input$county_name)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.subtitle = element_text(size=14, color = "grey20", face = "bold"))
  
g2/g3
})

```


